# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
This assumes that the file has already been downloaded and present in the current working directory
Now unzip it
```{r}
unzip("repdata.zip")
```
And read the data and convert the strings in the "date" variable to date
```{r}
repdata <- read.csv("activity.csv", stringsAsFactors=FALSE)
repdata$date <- as.Date(repdata$date, "%Y-%m-%d")
```
## What is mean total number of steps taken per day?
Show the histogram of daily number of steps
```{r}
hist(aggregate(steps ~ date, FUN=sum, data=repdata, na.rm=TRUE)$steps)
```
And calculate the mean and median of total daily steps (as single values summarizing the entire data set)
```{r}
mean(aggregate(steps ~ date, FUN=sum, data=repdata, na.rm=TRUE)$steps, na.rm=TRUE)
median(aggregate(steps ~ date, FUN=sum, data=repdata, na.rm=TRUE)$steps, na.rm=TRUE)
```
## What is the average daily activity pattern?
Make a time series line plot
```{r}
plot(aggregate(steps ~ interval, FUN=mean, data=repdata, na.rm=TRUE), type="l")
```
And show which interval has the maximum number of steps - it turns out to be interval 835
```{r}
temp.data <- aggregate(steps ~ interval, FUN=mean, data=repdata, na.rm=TRUE)
temp.data[which.max(temp.data$steps),]
```
## Imputing missing values
Report the number of missing values in the data set - there seem to be 2304 NAs
```{r}
sum(is.na(repdata$steps))
```
For the missing data imputation, I will use interval averages, because there are 8 dates with all NAs, but all intervals have 53 observations
```{r}
aggregate(!is.na(steps) ~ date, FUN=sum, data=repdata)
aggregate(!is.na(steps) ~ interval, FUN=sum, data=repdata)
```
The imputation code
```{r}
intavg <- aggregate(steps ~ interval, FUN=mean, data=repdata, na.rm=TRUE)
rownames(intavg) <- paste0("I", as.character(intavg$interval))
impdata <- repdata
for (i in 1:nrow(repdata)) {
  if (is.na(repdata[i, "steps"])){
    impdata[i, "steps"] <- intavg[paste0("I", as.character(repdata[i, "interval"])), 2]
  }
}
```
And repeat some previous steps after imputation
```{r}
hist(aggregate(steps ~ date, FUN=sum, data=impdata, na.rm=TRUE)$steps)
mean(aggregate(steps ~ date, FUN=sum, data=impdata, na.rm=TRUE)$steps, na.rm=TRUE)
median(aggregate(steps ~ date, FUN=sum, data=impdata, na.rm=TRUE)$steps, na.rm=TRUE)
```
The mean has not changed (due to the way imputation was performed), but the median changed due to the fact we now have more values 
## Are there differences in activity patterns between weekdays and weekends?
Create new factor variable - note that in this case, I used Polish names as this is the default on my system. That is why you might not recognize the week names
```{r}
impdata$daytype <- weekdays(impdata$date) %in% c("sobota", "niedziela")
impdata$daytype[impdata$daytype] <- "weekend"
impdata$daytype["FALSE" == impdata$daytype] <- "weekday"
impdata$daytype <- as.factor(impdata$daytype)
```
Create the plot
```{r}
par(mfrow=c(2,1))
with(impdata["weekend" == impdata$daytype,], plot(aggregate(steps ~ interval, FUN=mean), type="l", main="weekend"))
with(impdata["weekday" == impdata$daytype,], plot(aggregate(steps ~ interval, FUN=mean), type="l", main="weekday"))
```